# Introduction {#sec-intro}

```{r}
#| echo: false
source("_common.R")
status("drafting")
```

Data science is the discipline of turning raw data into understanding,
insight, and knowledge. The goal of this workshop is to give you a
practical, hands-on introduction to doing data science with R, in the
context of the **3XG study at VITO**.

## What you will learn

This workshop focuses on the essential first steps of any data analysis
project:

1. **Import** — getting your data into R, whatever format it originally
   comes in (CSV, Excel, etc.).
2. **Tidy** — reshaping data so each variable is a column, each
   observation is a row, and each value is a cell.
3. **Explore** — repeatedly transforming, visualizing, and modelling data
   to understand it better.
4. **Communicate** — presenting your results clearly to others.

Surrounding these steps is **programming**: the set of skills you use to
automate common tasks and to write clear, reusable code.

```{r}
#| label: fig-ds-diagram
#| echo: false
#| fig-cap: >
#|   A simplified data science workflow: import, tidy, and then iterate
#|   between transforming, visualizing, and modelling your data, before
#|   communicating the results.
#| fig-alt: >
#|   A diagram displaying the data science cycle: Import -> Tidy -> Understand
#|   (which contains Transform -> Visualize -> Model in a cycle) -> Communicate.
knitr::include_graphics("images/data-science-workflow.svg", error = FALSE)
```

## Why R?

R is a free, open-source programming language and environment designed
specifically for statistical computing and data analysis. It is widely
used across research institutes, including VITO, because it:

- Has a large ecosystem of packages for data wrangling, visualization,
  statistics, and reporting (many of them part of the
  [tidyverse](https://www.tidyverse.org/)).
- Makes reproducible analysis straightforward, which is essential for
  research studies like 3XG.
- Is free and works on Windows, macOS, and Linux.

## Tools you'll need

To follow along with this workshop, you'll need:

- **R** — download it from [CRAN](https://cran.r-project.org/).
- **RStudio** (or another R-friendly editor such as Positron or VS Code
  with the R extension) — download RStudio Desktop from
  [posit.co](https://posit.co/download/rstudio-desktop/).
- **Quarto** — used to build this workshop book and to write your own
  reports; see [quarto.org](https://quarto.org/).
- The **tidyverse** package, which bundles together the R packages used
  throughout this workshop:

```{r}
#| eval: false
install.packages("tidyverse")
```

## How this workshop is organized

After this introduction, the workshop covers:

- @sec-workflow-basics: the basics of writing and running R code.
- @sec-data-import: how to bring data into R so you can start exploring
  it, using an example dataset in the style of the data typically used in
  the 3XG study.

More chapters — covering data transformation, visualization, and
reporting — will be added as the workshop develops.

## Acknowledgements

This introduction is inspired by the structure of the introduction chapter
in [*R for Data Science*](https://r4ds.hadley.nz/intro.html), rewritten and
adapted for the 3XG/VITO context.
